---
title: "Wage Visualisations"
author: "Sookthi"
date: "2023-10-13"
output: html_document
---

### Load Libraries
```{r, message = FALSE}
library(tidyverse)
library(dbplyr)
library(readxl)
library(ggplot2)
```

### Load database
```{r, message = FALSE}
my_data <- read_excel("CRAN_Wage.xlsx")
```

### Create histogram 
```{r}
ggplot(data = my_data, mapping = aes(x = wage))+
  geom_histogram(binwidth=5)
```

### Summarise data 
```{r}

my_data %>%
  group_by(education) %>%
  summarise(mean(wage))

my_data %>%
  group_by(education) %>%
  summarise(median(wage))

```

### Scatter plot
Start with the CRAN_Wages data frame, and map the x axis to wage and y axis to age. 

Represent each observation with a point, corresponding to the age and wage of a recorded individual. The colour of the point will depend on the sector the individual is working in. 

`labs` is used to title the graph 'Wages and Ages', assign a subtitle to the graph, name the x and y axes, and attribute the data to a source in the caption. 

`scale_colour_viridis_d()` is used to make the colour of the points in the graph more accessible to colour-blind individuals. 

```{r}
ggplot(data=my_data, mapping = aes (x = wage, y = age, colour=jobclass))+
  geom_point()+
  labs(title = "Wages and Ages", 
       subtitle = "Dimensions across Industrial and Information jobs",
       x = "Wage",
       y = "Age",
       colour = "Sector",
       caption = "Source: CRAN_Wage") +
  scale_colour_viridis_d()
       
```



### Box and whisker plot


